**The Heart can Lie: The Role of Interoception and Theory of Mind in Deception**

### 2023-01-23 paper/manuscript.Rmd
edit manu method and procedures

The sequence of each trial was the same for both conditions. A question was shown on a computer screen, either phrased directly (e.g., "What is your favourite sport?"), or indirectly (e.g., "Is your favourite sport Hockey?"), with accompanying instructions below that either asked participants to verbally lie or tell the truth (half of the trials in each condition, i.e., n = 20, were directed lies). Following a short interval (0.7 - 1.5s) for response preparation (during which "Connecting..." was displayed on the screen), participants were presented with either a social-feedback or biofeedback cue (for the *Interrogation* and *Polygraph* conditions, respectively), displayed for a maximum of 10s. After the feedback cue, participants had to provide their answer to the question (verbally) aloud and press the 'space' key to mark the end of their response. Following another short interval (1.5 - 2.5s) during which was written "Disconnecting..." on the screen, participants were asked to rate, using an analog scale, their confidence in how convincing (i.e., likely to get judged as truthful) they perceived their response to be.

In the *Interrogation* condition, participants had to provide their answer each trial while receiving social feedback in the form of a video stream of an examiner, and were informed that the examiner would be evaluating the truthfulness of their responses after observing them through the live video feed. In actuality, the displayed video feed was pre-recorded video clips of a confederate's face (staying still with minimal reactions, and with a medical mask, as the experiment was run during COVID restrictions), and the same video-clips were used with all participants. Note that after debriefing, all participants reported believing that the stream was real and that the examiner was really there.

In the *Polygraph* condition, participants had to provide their answer answer each trial while receiving biofeedback in the form of physiological signals (ECG, EDA and Respiration) being recorded. Participants were informed that they were being shown live feedback of their own physiological activity. In actuality, we displayed pre-recorded video clips of a confederate's physiological signals. Note that after debriefing, all participants reported believing that it was indeed their own signals being shown.

@@

In the first phase of the study, participants answered a brief demographic survey before being administered the questionnaires and cognitive-behavioural tasks in a randomized order. Then, the physiological recording devices were set up as follows: Cardiac activity (ECG) was recorded with 3 electrodes placed according to a modified Lead II configuration [@takuma1995alternative]; electrodermal activity (EDA) and respiration (using a respiration belt) were also recorded, but not analyzed for the study (unfortunately, most participants did not yield any skin conductance responses - which we believe was partly caused by the low temperature (22 degrees Celsius) of the experimental room). All signals were recorded at 1000 Hz via the BioPac MP160 system (BioPac Systems Inc., USA). The deception task was finally run, followed by a debriefing. Three outcome variables were recorded for each trial of the deception task, namely the participants' confidence ratings that their answers (lies *vs.* truths) were convincing, the response time (RT) between the question onset and the participant's key press (indicating the end of their verbal answer), and the heart rate change associated with the response (within a window of 3.5 s).

@@

>>>>

The sequence of each trial was the same for both conditions. Participants were instructed to respond to a question shown on a computer screen, phrased either directly (e.g., "What is your favourite sport?"), or indirectly (e.g., "Is your favourite sport Hockey?"), by verbally lying or telling the truth (for half of the trials in each condition, i.e., n = 20). Following a short interval (0.7 - 1.5s) for response preparation, during which "Connecting..." was shown on the screen, a social or bio-feedback cue (for *Interrogation* and *Polygraph* conditions respectively) was displayed for a maximum of 10s. Participants then had to provide their answers verbally, and pressed the space key to signify the end of their response. After another short interval (1.5-2.5s), during which "Disconnecting" was written on the screen, participants were asked to rate their confidence in how convincing (i.e., likely to get judged as truthful) they perceived their response to be on a visual analog scale.

In the *Interrogation* condition, participants had to provide their answer while receiving social feedback in the form of a video stream of an examiner, and were informed that the examiner would be evaluating the truthfulness of their responses after observing them through the live video feed. In actuality, the displayed video feed were pre-recorded video clips of a confederate's face (staying still with minimal reactions, and with a medical mask, as the experiment was run during COVID restrictions), and the same video-clips were used with all participants. All participants reported believing that the stream was real and that the examiner was really there during the debriefing.

In the *Polygraph* condition, participants had to provide their answer while receiving biofeedback in the form of physiological signals (ECG, EDA and Respiration) being recorded. Participants were informed that they were being shown live feedback of their own physiological activity. In actuality, we displayed pre-recorded video clips of a confederate's physiological signals. All participants reported believing that it was indeed their own signals being shown during the debriefing.

@@

This study is comprised of 2 sessions. During session 1, participants answered a brief demographic survey before the questionnaires, which were presented in a randomized order. During session 2, cognitive-behavioural tasks were administered to participants while their physiological signals (including cardiac activity (ECG), respiration (RSP) and electrodermal activity (EDA)) were being recorded. The physiological recording devices were set up as follows: ECG was recorded with 3 electrodes placed according to a modified Lead II configuration [@takuma1995alternative], and respiration was measured using a respiration belt. All signals were recorded at 1000Hz via the BioPac MP160 system (BioPac Systems Inc., USA). For all participants, session 2 began with the deception task, followed by the YONI task and the HCT, with the latter two presented in a randomized order. Three outcome variables were recorded for each trial of the deception task, namely the participants' confidence ratings that their answers (lies *vs.* truths) were convincing, the response time (RT) between the question onset and the participant's key press (indicating the end of their verbal answer), and the heart rate change associated with the response (within a window of 3.5 s).

@@

EDA was not further analyzed for the study because most participants did not yield any skin conductance responses - which we believe was partly caused by the low temperature (22 degrees Celsius) of the experimental room.
